---
title: "Getting used to dictionaries"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

```{r, include = FALSE}
library("quanteda")
```


### Creating a simple dictionary

Dictionaries are named lists, consisting of a "key" and a set of entries defining
the equivalence class for the given key.  To create a simple dictionary of parts of
speech, for instance we could define a dictionary consisting of articles and conjunctions, using the `dictionary()` constructor

```{r}
posDict <- dictionary(list(articles = c("the", "a", "an"),
                           conjunctions = c("and", "but", "or", "nor", "for", "yet", "so")))
```

You can examine this dictionary by invoking its `print` method, simply by typing the name of the object and pressing Enter. Please try that now.

```{r}
posDict
```

Create a DFM called `posDfm` with the `data_corpus_inaugural` corpus (which comes with quanteda) and apply the dictionary. Print the first couple of rows of the DFM. Hint: `?head()`

```{r}
posDfm <- dfm(data_corpus_inaugural, dictionary = posDict)
head(posDfm)
```

Now, plot the values of articles and conjunction over the time (across the speeches). Make sure the x-axis shows the years.

Is the distribution of articles and conjunctions relatively constant across years, as you would expect?

**Your answer here**

> No, there is a lot of variation. The reason for that is that the raw number of articles and conjunctions will be a function of document length.

```{r}
plot(x = docvars(data_corpus_inaugural, "Year"), 
     y = posDfm[, "articles"],
     type = "p", pch = 16, col = "orange",
     ylim = range(posDfm),
     xlab = "Year", ylab = "Term frequency")
points(x = docvars(data_corpus_inaugural, "Year"), 
     y = posDfm[, "conjunctions"],
     pch = 3, col = "blue", new = FALSE)
```

Replicate the graph, but this time using weights in the `posDfm` that take into account the differences in the length of each speech. Do you find the same pattern? Interpret the pattern you observe.

**Your answer here**

> The graphic shows that the usage of articles vs conjunctions is not stable over time. During the 19th century the usage is more or less stable but then the relative importance of conjunctions compared to articles increases more and more. Since the 1990s both types of function words seem to be used about equally. This probably has to do with the fact that recent speeches have been shown on TV, with more frequent pauses for applause, which means presidents probably prefer to speak in shorter, simpler sentences.

```{r}
posDfmWeight <- dfm_weight(posDfm, scheme = "prop")
head(posDfmWeight)

# base R plot
plot(x = docvars(data_corpus_inaugural, "Year"), 
     y = posDfmWeight[, "articles"],
     type = "p", pch = 16, col = "orange",
     ylim = range(posDfmWeight),
     xlab = "Year", ylab = "Relative term frequency")
points(x = docvars(data_corpus_inaugural, "Year"), 
       y = posDfmWeight[, "conjunctions"],
       pch = 3, col = "blue", new = FALSE)

```

#### Hierarchical dictionaries

Dictionaries may also be hierarchical, where a top-level key can consist of subordinate keys, each a list of its own.  For instance, `list(articles = list(definite="the", indefinite=c("a", "an"))` defines a valid list for articles.  Make a dictionary of articles and conjunctions where you define two levels, one for definite and indefinite articles, and one for coordinating and  subordinating conjunctions.  (A sufficient list for your purposes of  subordinating conjuctions is "although", "because", "since", "unless".)

```{r}
posDictHier <- list(
  article = list(definite = "the", indefinite = c("a", "an")),
  conjunction = list(
    coordinating = c("and", "but", "or", "nor", "for", "yet", "so"),
    subordinating = c("although", "because", "since", "unless")
  )
)
```

Now apply this to the `data_corpus_inaugural` object, and examine the resulting features. 

```{r}
posDfmHier <- dfm(data_corpus_inaugural, dictionary = dictionary(posDictHier))
head(posDfmHier)
```


Do the subcategories of the DFM created with `posDictHier` sum to the two general categories created with `posDict`? Write code to check if that's the case. Briefly explain why/why not.

**Your answer here**

>For the article category they do but not for the conjunctions. This is to be expected as in the first version (`posDfm`) only coordinating conjunctions were included whereas the hierarchical categories also contains subordinating conjunctions.

```{r}
posDfmHierAlt <- dfm_lookup(posDfmHier, dictionary = dictionary(list(
  article = c("article.definite", "article.indefinite"), 
  conjunction = c("conjunction.coordinating", "conjunction.subordinating")
)))
head(posDfmHierAlt)
head(posDfm)

# there are more elements in posDictHier than in posDict
length(unlist(posDictHier))
length(unlist(posDict))
```


        
### Getting used to thesauruses 

A "thesaurus" is a list of feature equivalencies specified in the same list format as a dictionary, but with a key difference. Which one? Let's find out.

If we wanted to count pronouns as equivalent, for instance, we could use the thesaurus argument to `dfm` in order to group all listed prounouns into a single feature labelled "PRONOUN".

```{r}
mytexts <- c("We are not schizophrenic, but I am.", "I bought myself a new car.")
myThes <- dictionary(list(pronouns = list(firstp = c("I", "me", "my", "mine", "myself", "we", "us", "our", "ours"))))
myDfm <- dfm(mytexts, thesaurus = myThes)
myDfm
```

Notice how the thesaurus key has been made into uppercase - this is to identify it as a key, as opposed to a word feature from the original text.



```{r}
myDfmFunc <- dfm(mytexts, thesaurus = posDict)
myDfmFunc
```

When you call `dfm()` with a `dictionary = ` or `thesaurus = ` argument, then what `dfm()` does internally is actually to first construct the entire dfm, and then select features using a call to `dfm_lookup()`.

Try creating a dfm object using a sample of five inaugural speeches, with no dictionary applied. Then apply the `posDict` from the earlier question to select features 

a) in a way that replicates the `dictionary` argument to `dfm()`, 
b) in a way that replicates the `thesaurus` argument to `dfm()`.

```{r}
set.seed(123)
inaugFive <- corpus_sample(data_corpus_inaugural, 5)
inaugFiveDfm <- dfm(inaugFive)

# a) Replicate "dictionary = posDict"
dfm_lookup(inaugFiveDfm, posDict, exclusive = TRUE)

# b) Replicate "thesaurus = posDict"
dfm_lookup(inaugFiveDfm, posDict, exclusive = FALSE)
```


