---
title: "Getting used to dictionaries"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

```{r, include = FALSE}
library("quanteda")
```


### Creating a simple dictionary

Dictionaries are named lists, consisting of a "key" and a set of entries defining
the equivalence class for the given key.  To create a simple dictionary of parts of
speech, for instance we could define a dictionary consisting of articles and conjunctions, using the `dictionary()` constructor

```{r}
posDict <- dictionary(list(articles = c("the", "a", "an"),
                           conjunctions = c("and", "but", "or", "nor", "for", "yet", "so")))
```

You can examine this dictionary by invoking its `print` method, simply by typing the name of the object and pressing Enter. Please try that now.

```{r}

```

Create a DFM called `posDfm` with the `data_corpus_inaugural` corpus (which comes with quanteda) and apply the dictionary. Print the first couple of rows of the DFM. Hint: `?head()`

```{r}

```

Now, plot the values of articles and conjunction over the time (across the speeches). Make sure the x-axis shows the years.

Is the distribution of articles and conjunctions relatively constant across years, as you would expect?

**Your answer here**

> 

```{r}
plot(x = docvars(data_corpus_inaugural, "Year"), 
     y = posDfm[, "articles"],
     type = "p", pch = 16, col = "orange",
     ylim = range(posDfm),
     xlab = "Year", ylab = "Term frequency")
points(x = docvars(data_corpus_inaugural, "Year"), 
     y = posDfm[, "conjunctions"],
     pch = 3, col = "blue", new = FALSE)
```


Replicate the graph, but this time using weights in the `posDfm` that take into account the differences in the length of each speech. Do you find the same pattern? Interpret the pattern you observe.

**Your answer here**

> 

```{r}

```

#### Hierarchical dictionaries

Dictionaries may also be hierarchical, where a top-level key can consist of subordinate keys, each a list of its own.  For instance, `list(articles = list(definite="the", indefinite=c("a", "an"))` defines a valid list for articles.  Make a dictionary of articles and conjunctions where you define two levels, one for definite and indefinite articles, and one for coordinating and  subordinating conjunctions. (A sufficient list for your purposes of  subordinating conjuctions is "although", "because", "since", "unless".)

```{r}
posDictHier <- list(
  article = list(definite = "the", indefinite = c("a", "an")),
  conjunction = list(
    coordinating = c("and", "but", "or", "nor", "for", "yet", "so"),
    subordinating = c("although", "because", "since", "unless")
  )
)
```

Now apply this to the `data_corpus_inaugural` object, and examine the resulting features. 

```{r}

```



Do the subcategories of the DFM created with `posDictHier` sum to the two general categories created with `posDict`? Write code to check if that's the case. Briefly explain why/why not.

**Your answer here**

>

```{r}

```



        
### Getting used to thesauruses 

A "thesaurus" is a list of feature equivalencies specified in the same list format as a dictionary, but with a key difference. Which one? Let's find out.

If we wanted to count pronouns as equivalent, for instance, we could use the thesaurus argument to `dfm` in order to group all listed prounouns into a single feature labelled "PRONOUN".

```{r}
mytexts <- c("We are not schizophrenic, but I am.", "I bought myself a new car.")
myThes <- dictionary(list(pronouns = list(firstp = c("I", "me", "my", "mine", "myself", "we", "us", "our", "ours"))))
myDfm <- dfm(mytexts, thesaurus = myThes)
myDfm
```

Notice how the thesaurus key has been made into uppercase - this is to identify it as a key, as opposed to a word feature from the original text.


```{r}
myDfmFunc <- dfm(mytexts, thesaurus = posDict)
myDfmFunc
```

When you call `dfm()` with a `dictionary = ` or `thesaurus = ` argument, then what `dfm()` does internally is actually to first construct the entire dfm, and then select features using a call to `dfm_lookup()`.

Try creating a dfm object using a sample of five inaugural speeches, with no dictionary applied. Then apply the `posDict` from the earlier question to select features 

a) in a way that replicates the `dictionary` argument to `dfm()`, 
b) in a way that replicates the `thesaurus` argument to `dfm()`. 

```{r}
set.seed(123)

```


