---
title: 'CPS Formative Assignment Part B \n Solutions'
subtitle: 'Data exploration'
author: "YOUR NAME HERE"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

```{r, include = F}
library("quanteda")
library("stringr")
```

### Intro

In this assignment we continue our work on data exploration using Twitter data. Among other things, we employ regular expressions to compute some useful descriptive statistics for this type of data.  


### Question 1 (1 point)

The file `candidate-tweets.csv` that contains slightly over 25,000 tweets published by the 4 leading candidates in the 2016 Presidential Primary Elections in the US (Cruz, Trump, Clinton, Sanders). 

Read the dataset into R using `read.csv()`. It is recommended to set the argument `stringsAsFactors=FALSE` to assure that text data is parsed as the correct data type: as characters and not as factors. Make sure that the file `candidate-tweets.csv` is in the same folder as the `.Rmd` file.

Use the functions `head()` and `str()` to inspect the structure of the data. 

```{r}

```


### Question 2 (1 point)

About how many tweets are retweets? What is the percentage of retweets among all tweets? Use regular expressions to answer these questions.

**Your answer here**

>

```{r}

```

### Question 3 (1 point)

How many tweets were sent by each of the candidates? Hint: `?table`

```{r}

```




### Question 4 (3 points)

Create a (smaller) data frame that contains only tweets by Bernie Sanders. Hint: You might want to create vector that indicates the rows that hold tweets from Bernie Sanders and use this vector to create a smaller data frame.

```{r}

```

How many times did Bernie Sanders mention words related to 'health care'? and 'immigration'? and 'billionaires'? 'gun control'? 'poverty'? Use regular expressions to make sure your searches return all relevant results.

**Your answer here**

>

```{r}

```


### Question 5 (5 points)

What are the 10 most frequent hashtags in the tweets by Bernie Sanders? Try to answer this question creating a function that extracts hashtags and sorts them by their frequency in decreasing order. 


```{r}

```

Now, find the 10 most frequent hashtags for each candidate. You can use a loop to answer this question or just run the above function for each candidate separately.

```{r}

```



### Question 6 (3 points)

Going back to the dataset with tweets by Bernie Sanders, try to create a corpus and a document feature matrix. Think carefully through the different options (e.g. should you exclude stopwords?). Then, look at the 25 most frequent words. What do you learn?

**Your answer here**

>

```{r}

```

Plot the DFM you just created using a wordcloud. Do you learn more/same/less about the data? 

**Your answer here**

>

```{r}

```

### Question 7 (2 points)

Regular expressions are very important concepts in text processing, as they offer
tools for searching and matching text strings based on symbolic representations.
For the dictionary and thesaurus features, we can define equivalency classes in terms of regular expressions (see tutorial on regular expressions at <http://www.regular-expressions.info>).

This provides an easy way to recover syntactic variations on specific words, without relying on a stemmer. For instance, we could query a regular expression on tax-related words, using:

```{r, eval = FALSE}
kwic(data_corpus_inaugural, "tax", valuetype = "regex")
kwic(data_corpus_inaugural, "tax")
```

Why is the result different than the output you get when you run `kwic(data_corpus_inaugural, "tax")`?

**Your answer here**

>

What if we on wanted to construct a regular expression to query only "valued" and "values" but not other variations of the lemma "value"?

```{r}

```
